Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals

### 2021-09-30 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex

% An example for the ar2rc document class.
% Copyright (C) 2017 Martin Schroen
% Modifications Copyright (C) 2020 Kaishuo Zhang
%
% This program is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
%
% This program is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program.  If not, see <http://www.gnu.org/licenses/>.

% Use quote enviromnent for manuscript text

\documentclass{AR2RC}

\usepackage{siunitx}

\title{Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals}
\author{Mathieu Simon, Michael Indermaur, Denis Schenk, Seyedmahdi Hosseinitabatabaei, Bettina	M. Willie and Philippe Zysset}
\journal{Bone}
\doi{BONE-D-21-00758}

\begin{document}

\maketitle

\vspace{1em}The authors would like to thanks the reviewers for their time spent for this article review. The work they performed will allow to improve the quality of the present article and help the authors to transmit the knowledge resulting of their work with more clarity. For this, we answered at your comments as presented in this document.

\section{Reviewer \#1}

\RC This report analyzed three dimensional morphology and stiffness of trabecular bone in patients with Osteogenesis Imperfecta (OI), and compared them with those from healthy control group. 3D morphologies were extracted from HR-pQCT data; while stiffness information was obtained by numerical simulation techniques. This is an important research topic with potential applications in assessing bone fragility as a result of OI. However, this reviewer felt the manuscript did not provide sufficient information to clearly explain new progress and novelty.

\AR To explain new progress shown in the present article, the manuscript was modified at multiple places as follow:\par

Abstract, Line 23-24
\begin{quote}
	Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls \Add{because validity of these relationships remains unclear for OI bone and are crucial for bone strength assessment using hFE}.
\end{quote}

Abstract, Line 48-49
\begin{quote}
	 Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections. \Add{Thus, the hFE scheme used for bone strength assessment of healthy individuals is valid for OI individuals as well.}
\end{quote}

\subsection{Methods}

\RC The general methodology used in this paper followed closely a previously published paper by the same senior author (Zysset, ref 12) on human trabecular bone. Briefly, CT imaging of trabecular bone was followed by micro-FE modeling for stiffness (FE stiffness), and a fabric-elasticity model (Predicted stiffness) reported also by the same senior author (Zysset) in 1995 (ref 32 ). Other than applications of the protocols to OI patients, contributions to either imaging or the modeling techniques seem to be moderate.

\AR You have raised an important point here. The methodology used in the present paper is based on previous publications for trabecular bone analysis. With time, these different works led to the development of the hFE scheme for bone strength estimation. However, the work cited here (Zysset, ref 12), is based on \si{\micro}CT scan and was performed on human vertebrae, proximal femur and distal radius. It is the case for another previous work (Zysset, ref 24) where an analysis of the influence of boundary conditions type was performed. Again, stiffness tensor was computed from numerical simulations of trabecular bone from a \si{\micro}CT scans on the proximal femur. The present study extend the current knowledge by 1) applying this hFE methodology at the distal tibia 2) using HR-pQCT scans i.e. lower resolution scans 3) with an important control cohort of 120 healthy individuals and 4) demonstrating the applicability of the hFE scheme on OI diagnosed individuals. More emphasis on these points is put with the following modifications of the manuscript:\par

Introduction, Line 44-45
\begin{quote}
	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
\end{quote}

\subsection{Article contribution}

\RC Key progress and contribution are in the conclusion that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, based on which one could use HR-pQCT based hFE outcomes for fracture risk assessment. This conclusion was made based on filtered data (Figure 7), and therefore is valid only when trabecular bone is homogeneous enough. When all data including those with low BV/TV are included, the correlation between the fitted stiffness tensor and the observed one is different between the OI group and the healthy group. In other words, the conclusion is no longer valid for trabecular bone with low BV/TV. In terms of fracture risk, trabecular bone with lower density (low BV/TV) in OI patients would pose higher fracture risk, but could not be accurately predicted with the proposed method. Therefore, progress related to OI bone seems to be limited.

\AR Explain filtering of low stiffness means filtering of low BV/TV...

\subsection{OI brittelness}

\RC OI affects the production of collagen Type I, and therefore the brittleness of the bone matrix (a composite of bone mineral and collagen protein). Current methodology seemed to be limited to elasticity theory. Its application to predicting bone fragility (toughness) in the case of OI needs has yet to be proved.

\AR Michi articles

\section{Reviewer \#2}

\RC The manuscript showed a mathematical investigation of the fabric-elasticity relationships in case of osteogenesis imperfecta and healthy trabecular bones, obtained from distal tibiae. The evidence suggested that the fabric-elasticity relationships of sufficiently homogeneous trabecular tissue with osteogenesis imperfecta and healthy trabecular tissue are similar.

\subsection{Clinical output}

\RC Methodologically, the paper is interesting and technically sound. It is clear and easy to follow. I do appreciate the direct comparison with the literature, and the sample size collected by the authors. On the other hand, as a Bone reader, I have found unusual reading this kind of engineering/mathematical paper here, and I suppose it is far from the typical audience of the journal. I understand there is a final clinical message, but the way it is reached could be misleading. I would have expected to read this kind of paper in a more engineering journal. I suggest to increase and clarify the possible clinical output of the study in order to increase the impact.

\AR Clarifications in text

\subsection{OI type differentiation}

\RC Another concern consists in the grouping of the bones with OI. Indeed, while the OI group was composed of patients with a diagnosis of OI Type I, III, and IV, the tibiae were merged in a single group (OI group), without any differentiation during the following steps of the study. In the literature and in the studies cited by the authors, the differences among bones with different types of OI resulted remarkable. Thus, I would like to find more details about the OI types in the linear regressions (differentiating the regressions for each type of OI type). Moreover, it would be interesting to see which is the OI type more influenced by the filtering. My idea is that bones with OI type I are "similar" to healthy group, they can easily respect the homogeneity condition, therefore they were not excluded by the filtering strategies. By contrast, the type IV may have a microstructure different from the healthy bones, and they can explain the lack of regression in the OI group. If this is the case, the final message should be mitigated.

\AR Change figures in linear regressions to highlight the similarities among the OI types

\subsection{Introduction}
Introduction: please, add statistics also about the epidemiology of OI in men and women. Moreover, add the lacks in the literature about the OI that this study could solve.

\subsection{Methods - Participants}
Could you add statistics about body anthropometry? I was wondering, why the groups were matched considering age and not height and weight? Could you explain your choice?

\subsection{Methods - HR-pQCT}
The authors used two different protocols to identify the region of interest in the tibiae. Considering that the same protocol could be applied on both groups (control and OI) I do not understand why two different protocols were used. There is a difficult in defining the dense structure in OI specimens? To clarify the manuscript, I suggest to label the most proximal stack in the control tibia as Stack n. 1

\subsection{Methods - Image analysis}
Pag 3, lines 34-35, left column. The stack was divided into two halves WITH A TRANSVERSE PLANE. Pag 3, lines 4-7, right column. Did you use any filter to reduce the noise? If yes, please specify the procedure. Pag 4, line 4. Does your concern about the rotation consist in the errors associated with interpolation?

\subsection{Methods - Linear Regression}
I would specify here that you are considering the entire sample.

\subsection{Methods - Data filtering}
You removed the regions that violated the assumption of homogeneity. Could this affect the final findings of the study? I mean, could this filtering remove the specimens with OI type IV?

\subsection{Resutlts - Morphological analysis}
Please, report here the distribution of OI types in the cases matched with controls.

\subsection{Table 2}
Make easier and more effective the comparison with the literature, indicating below the table the OI types analyzed in the other studies. The positioning on different lines, in the same cell, could not be caught.

\subsection{Results - Comparison with literature}
Could you perform a test to evaluate if differences exist among the different regressions?

\subsection{Discussion}
The discussion is interesting and full of details and rigorous considerations. However, it is based only on the engineering/mathematical approach used in the study, and far from the clinical audience. I would like to read more about the effects of the different OI types in the provided findings.

\end{document}

### 2021-10-01 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex

\RC The general methodology used in this paper followed closely a previously published paper by the same senior author (Zysset, ref 12) on human trabecular bone. Briefly, CT imaging of trabecular bone was followed by micro-FE modeling for stiffness (FE stiffness), and a fabric-elasticity model (Predicted stiffness) reported also by the same senior author (Zysset) in 1995 (ref 32 ). Other than applications of the protocols to OI patients, contributions to either imaging or the modeling techniques seem to be moderate.

\AR You have raised an important point here. The methodology used in the present paper is based on previous publications for trabecular bone analysis. With time, these different works led to the development of the hFE scheme for bone strength estimation. However, the work cited here (Zysset, ref 12), is based on \si{\micro}CT scan and was performed on human vertebrae, proximal femur and distal radius. It is the case for another previous work (Zysset, ref 24) where an analysis of the influence of boundary conditions type was performed. Again, stiffness tensor was computed from numerical simulations of trabecular bone from a \si{\micro}CT scans on the proximal femur. The present study extend the current knowledge by 1) applying this hFE methodology at the distal tibia 2) using HR-pQCT scans i.e. lower resolution scans 3) with an important control cohort of 120 healthy individuals and 4) demonstrating the applicability of the hFE scheme on OI diagnosed individuals. More emphasis on these points is put with the following modifications of the manuscript:\par

Introduction, Line 44-45
\begin{quote}
	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
\end{quote}

Methods - HR-pQCT, Line 28-29
\begin{quote}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Brütisellen, Switzerland) was performed at the distal tibia on all patients from both groups. \Add{The distal tibia was chosen because numerous studies were performed on the distal radius [2, 5, 12, 14, 16, 31], some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) [1, 4, 9] but only one study with XtremeCTII was performed on the distal tibia [27]. Good bone strength assessment of tibia is crucial for individuals in wheelchair. As their tibiae do not sustain the body weight anymore, remodeling takes place and low energy fracture can occur.}
\end{quote}

### 2021-10-01 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex

Abstract, Lines 48-49
\begin{quote}
	 Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections. \Add{Thus, the hFE scheme used for bone strength assessment of healthy individuals is valid for OI individuals as well.}
\end{quote}

\subsection{Methods}

\RC The general methodology used in this paper followed closely a previously published paper by the same senior author (Zysset, ref 12) on human trabecular bone. Briefly, CT imaging of trabecular bone was followed by micro-FE modeling for stiffness (FE stiffness), and a fabric-elasticity model (Predicted stiffness) reported also by the same senior author (Zysset) in 1995 (ref 32 ). Other than applications of the protocols to OI patients, contributions to either imaging or the modeling techniques seem to be moderate.

\AR You have raised an important point here. The methodology used in the present paper is based on previous publications for trabecular bone analysis. With time, these different works led to the development of the hFE scheme for bone strength estimation. However, the work cited here (Zysset, ref 12), is based on \si{\micro}CT scan and was performed on human vertebrae, proximal femur and distal radius. It is the case for another previous work (Zysset, ref 24) where an analysis of the influence of boundary conditions type was performed. Again, stiffness tensor was computed from numerical simulations of trabecular bone from \si{\micro}CT scans on the proximal femur. The present study extend the current knowledge by 1) applying this hFE methodology at the distal tibia, 2) using HR-pQCT scans i.e. lower resolution scans than the two references cited here, 3) with an important control cohort of 120 healthy individuals and 4) demonstrating the applicability of the hFE scheme on OI diagnosed individuals. More emphasis on these points is put with the following modifications of the manuscript:\par

Introduction, Lines 44-45
\begin{quote}
	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
\end{quote}

Methods - HR-pQCT, Lines 28-29
\begin{quote}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Brütisellen, Switzerland) was performed at the distal tibia on all patients from both groups. \Add{The distal tibia was chosen because numerous studies were performed on the distal radius [2, 5, 12, 14, 16, 31], some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) [1, 4, 9] but only one study with XtremeCTII was performed on the distal tibia [27]. Accurate bone strength assessment of tibia is crucial for individuals in wheelchair. As their tibiae do not sustain the body weight anymore, remodeling takes place and low energy fracture can occur.}
\end{quote}

Results - Comparison with literature, Lines 61-62
\begin{quote}
	Table 4 shows results obtained \Add{using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m} compared to \Del{literature} \Add{previous femura studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m [12, 24]}.
\end{quote}

Discussion, Page 10, Lines 55-56, Left column
\begin{quote}
	To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively. \Add{These high numbers of participants allow to obtain strong results to compare then with literature.}
\end{quote}

Discussion, Page 10, Lines 38-40, Right column
\begin{quote}
	With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone. \Add{Thus, the same hFE analysis can be applied for bone strength assessment of OI diagnosed individuals or of healthy ones. Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity\Del{. Nevertheless}, the error created by such ROIs is negligible as this  concerns ROIs with extremely low BV/TV and therefore stiffness leading to a minor impact on the full model.
\end{quote}

Discussion, Page 11, Lines 39-43, Right column

\begin{quote}
	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e. with a CV lower than 0.263 and were matched for BV/TV and DA.\Add{Therefore, standard hFE analysis used for bone strength estimation of healthy individuals can be used for OI diagnosed individuals as well.}
\end{quote}

### 2021-10-03 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex

% An example for the ar2rc document class.
% Copyright (C) 2017 Martin Schroen
% Modifications Copyright (C) 2020 Kaishuo Zhang
%
% This program is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
%
% This program is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program.  If not, see <http://www.gnu.org/licenses/>.

% Use quote enviromnent for manuscript text

\documentclass{AR2RC}

\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{hyperref}

\hypersetup{hidelinks}

@@

\begin{quote}
	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e. with a CV lower than 0.263 and were matched for BV/TV and DA.\Add{Therefore, standard hFE analysis used for bone strength estimation of healthy individuals can be used for OI diagnosed individuals as well.}
\end{quote}

\subsection{Article contribution}

\RC Key progress and contribution are in the conclusion that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, based on which one could use HR-pQCT based hFE outcomes for fracture risk assessment. This conclusion was made based on filtered data (Figure 7), and therefore is valid only when trabecular bone is homogeneous enough. When all data including those with low BV/TV are included, the correlation between the fitted stiffness tensor and the observed one is different between the OI group and the healthy group. In other words, the conclusion is no longer valid for trabecular bone with low BV/TV. In terms of fracture risk, trabecular bone with lower density (low BV/TV) in OI patients would pose higher fracture risk, but could not be accurately predicted with the proposed method. Therefore, progress related to OI bone seems to be limited.

\AR Effectively, the key contribution, the fact that fabric-elasticity relationships in OI trabecular bone are similar to healthy bone, is based on filtered data. Nevertheless, the filtering was performed according to a coefficient of variation (CV) threshold, i.e. a homogeneity threshold. Figure 6, see Figure \ref{CV_BVTV}, shows CV in relation to BV/TV and it can be seen from this Figure that ROIs presenting a high heterogeneity have a lower BV/TV. Moreover, this tendency is confirmed by the Pearson's correlation coefficient being strictly different from zero, meaning there is a monotonic relation between CV and BV/TV. The fact that we want to highlight here is that a high CV means a low BV/TV and therefore, a low stiffness since it is mainly driven by BV/TV. This is confirmed by comparing Figure 5 and 7, see Figure \ref{CV_Filtering}. Both Figures have the same logarithmic scales and the filtering removes mainly the ROIs with a stiffness orders of magnitudes lower than the stiffness of homogeneous ROIs. Thus, when hFE is performed for bone strength assessment, ROIs that are mostly empty (low BV/TV) do not participate significantly (orders of magnitude lower) to the whole structure strength meaning that the error performed on such ROI stiffness can be neglected.

\begin{figure}[h!]
	\centering
	\begin{subfigure}[t]{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]
		{Pictures/03_CV_BVTV}
		\caption{CV in relation to BV/TV.}
		\label{CV_BVTV}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=\textwidth, trim= 250 140 250 0]
		{Pictures/R1_CVFiltering}
		\caption{Linear regression of the complete data set, healthy and OI pooled together. Shape differs according to CV value of the ROI and color according to the BV/TV.}
		\label{CV_Filtering}
	\end{subfigure}
	\caption{Relations between coefficient of variation (CV) and BV/TV}
	\label{01_Contribution}
\end{figure}

@@

\subsection{OI brittelness}

\RC OI affects the production of collagen Type I, and therefore the brittleness of the bone matrix (a composite of bone mineral and collagen protein). Current methodology seemed to be limited to elasticity theory. Its application to predicting bone fragility (toughness) in the case of OI needs has yet to be proved.

\AR Yes, OI is mainly characterized by impaired synthesis of type I collagen which compose the bone ECM and could therefore affects its toughness. To investigate this phenomenon, a first study was performed by Indermaur and al. (ref 15). They shown that the compressive strength of bone at the ECM level is related to degree of mineralization and that it is not reduced in OI compared to healthy individuals. A second study (not published yet) was performed to compare OI and healthy bone at the ECM level in tension. Their results shown no significant differences between the two phenotypes (i.e. healthy and OI). This, together with results of the present article, allow to conclude that fabric-strength relationships are similar between healthy and OI bone. This means that the main cause of bone fragility in OI is linked to the low bone mass.

@@

\RC The manuscript showed a mathematical investigation of the fabric-elasticity relationships in case of osteogenesis imperfecta and healthy trabecular bones, obtained from distal tibiae. The evidence suggested that the fabric-elasticity relationships of sufficiently homogeneous trabecular tissue with osteogenesis imperfecta and healthy trabecular tissue are similar.

\subsection{Clinical output}

\RC Methodologically, the paper is interesting and technically sound. It is clear and easy to follow. I do appreciate the direct comparison with the literature, and the sample size collected by the authors. On the other hand, as a Bone reader, I have found unusual reading this kind of engineering/mathematical paper here, and I suppose it is far from the typical audience of the journal. I understand there is a final clinical message, but the way it is reached could be misleading. I would have expected to read this kind of paper in a more engineering journal. I suggest to increase and clarify the possible clinical output of the study in order to increase the impact.

\AR Thank you for this remark. The main reason for the technical details was the clarity of the methods used to compare OI with healthy trabecular bone. On the other hand, Bone include articles with technical components as well. To clarify the clinical significance of this study, the manuscript is modified as follow:

### 2021-10-05 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Beginning of article revision
Response to reviewers draft following II

%\usepackage[authoryear,longnamesfirst]{natbib}
%\usepackage[authoryear]{natbib}
\usepackage[numbers]{natbib}
\usepackage{lipsum}
\usepackage{xcolor}
\usepackage[justification=centering]{caption}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{float}
\usepackage{multicol}

@@

@@

Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels. Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well. Bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls. Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different. At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior (ultimate stress and ultimate strain) than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\

High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}. Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008}. Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}. Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk. However, HR-pQCT-based hFE relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.

@@

In general, six cubic ROIs were selected at random position in each scan in a predefined area. Each ROI had to contain trabecular bone, but no cortical bone. For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 3) to be at roughly the same anatomical location as for the OI group. Then for both the healthy group and the OI group, the stack was divided into two halves and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}. For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone. This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\

@@

\section{Discussion}
Osteogenesis imperfecta is an inherited form of bone fra\-gility with a severity going from mild to perinatally lethal. This study aims to confirm that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment. To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively. \\

@@

The main limitations of this study are the definition of "homogeneous enough" (i.e. low CV) and the fact that it is limited to  XCTII reconstructions of the distal tibia. Moreover, having only one patient with OI type III where we could extract ROIs does not allows to perform statistics. %As those patients are less active, it could be interesting to analyse the impact of this condition on the weight bearing tibia. 
ROI homogeneity is a necessary condition underlying homogenisation of mechanical properties and the use of fabric-elasticity relationships. %has an important impact on the analysis quality. 
As proposed earlier, ROI homogeneity could be assessed with an alternative ROI filtering and more efforts could be invested to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROIs with extremely low stiffness, the impact on hFE models should remain negligible.
%if the proportion of such ROIs stays low.
A similar study could be performed on radius XCTII reconstructions to generalise the above findings. Another minor limitation is the slightly different protocol for selection of the anatomical site scanned in the control and OI groups. The section scanned for the control group was at a fixed distance, while the section scanned for the OI group was placed at a distance relative to the total length of the tibia. Nevertheless, as shown in Figure \ref{01_ClinicalSections}, the locations compare well and the ROIs were matched for BV/TV and DA to compare the fabric-elasticity relationships. A final limitation is that the measurements were performed on different HR-pQCT devices which were not cross-calibrated. However, as they are the same model (XCTII) and the quality control measurement ensures $\pm$ 1\% precision on BMD, this is expected to have a minor impact. \\

>>>>

%\usepackage[authoryear,longnamesfirst]{natbib}
%\usepackage[authoryear]{natbib}
\usepackage[numbers]{natbib}
\usepackage{lipsum}
\usepackage{xcolor}
\usepackage[justification=centering]{caption}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{float}
\usepackage{multicol}
\usepackage[normalem]{ulem}

@@

\newcommand{\Add}[1]{{\color{blue}{#1}}}

\newcommand{\Del}[1]{{\color{red}\sout{#1}}}

@@

	\Add{OI occurs worldwide and without sex preference. OI type I is estimated to occur in 45\% of the total OI cases, type II about 10\%, type III 25\% and type IV 20\% \cite{Martin2007}}.Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels. Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018} \Add{but aBMD is limited for bone strength estimation \cite{Pasco2006}}.  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well. Bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls. Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different. At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior (ultimate stress and ultimate strain) than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\

	High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}. Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008}. \Add{HR-pQCT-based FEA becomes more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}. Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk. However, HR-pQCT-based hFE relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.

@@

	In general, six cubic ROIs were selected at random position in each scan in a predefined area. Each ROI had to contain trabecular bone, but no cortical bone. For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number \Del{3}\Add{1}) to be at roughly the same anatomical location as for the OI group. Then, for both the healthy group and the OI group, the stack was divided into two halves \Add{with a transverse plane} and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}. For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone. This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\

@@

	\section{Discussion}
	Osteogenesis imperfecta is an inherited form of bone fra\-gility with a severity going from mild to perinatally lethal. This study aims to confirm that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment. To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively. \\

@@

	The main limitations of this study are the definition of "homogeneous enough" (i.e. low CV) and the fact that it is limited to  XCTII reconstructions of the distal tibia. Moreover, having only one patient with OI type III where we could extract ROIs does not allows to perform statistics. %As those patients are less active, it could be interesting to analyse the impact of this condition on the weight bearing tibia. 
	ROI homogeneity is a necessary condition underlying homogenisation of mechanical properties and the use of fabric-elasticity relationships. %has an important impact on the analysis quality. 
	As proposed earlier, ROI homogeneity could be assessed with an alternative ROI filtering and more efforts could be invested to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROIs with extremely low stiffness, the impact on hFE models should remain negligible.
	%if the proportion of such ROIs stays low.
	A similar study could be performed on radius XCTII reconstructions to generalise the above findings. Another minor limitation is the slightly different protocol for selection of the anatomical site scanned in the control and OI groups. The section scanned for the control group was at a fixed distance, while the section scanned for the OI group was placed at a distance relative to the total length of the tibia. Nevertheless, as shown in Figure \ref{01_ClinicalSections}, the locations compare well and the ROIs were matched for BV/TV and DA to compare the fabric-elasticity relationships. A final limitation is that the measurements were performed on different HR-pQCT devices which were not cross-calibrated. However, as they are the same model (XCTII) and the quality control measurement ensures $\pm$ 1\% precision on BMD, this is expected to have a minor impact. \\
